\documentclass{article}
    \usepackage[margin=1in]{geometry}
    \usepackage{amsmath,cancel}
    
    \title{CSCI 305, Homework \# 3}
    \author{Caleb Ouellette}
    \date{Due date:  Tuesday, May 8, midnight.}
    
    
    \begin{document}
    
    \maketitle
    
    \begin{itemize}
      \item
    Find $\Theta(T(n))$ for each case.
    \item
    For problem 1, find the solution three ways:
    \begin{itemize}
    \item The master theorem
    \item The substitution method
    \item Iterate, cancel, and solve the summation
      to find an exact solution.
    \end{itemize}
    \item
    For all other problems, use just the master theorem.
    \item
    When using the master theorem, be explicit about $n^{\log_b a}$,
    and about which case of the master theorem applies,
    about making sure the special condition applies in case 3
    of the master theorem, and
    about proving $f$'s correct
    asymptotic relationship to $n^{\log_b a}$.
    \end{itemize}
    
    \begin{enumerate}
    \item
    
      Solve this one three different ways:  master theorem,
      substitution method, and the iterate/cancel/summation method.
    
      \[T(n) = 2T(n/2) + n^4\]
      Master theorem \\
      \begin{eqnarray*}
      f(n) &vs.& n^{log_b a} \\
      n^4 &vs.& n^{log_2 2} \\
      n^4 &vs.& n^1 \\
      f(n) &=& \Omega(n^{log_b a+\epsilon}) \\
      T(n) &=& \Theta(n^4) \\
      \end{eqnarray*}
      note:
      $af(n/b) = 2f(n/2) = 4/2^4f(n) \leq cf(n)$
      \\
      \\
      Substitution method \\
      Guess $T(n) \leq c(n^4)$ \\
      \begin{eqnarray*}
      T(n) &=& 2T(n/2) + n^4 \\
      T(n) &\leq& 2c(n/2)^4 + n^4 \\      
      T(n) &\leq& c/8(n^4) + n^4 \\
      T(n) &\leq& (c/8 + 1)n^4 \\
      T(n) &\leq& c(n^4) \\
      \end{eqnarray*}
      for $c \leq 8/7$ \\
      We can use this same proof for $T(n) = \Omega(n^4)$ by flipping the signs, So $T(n) = \Theta(n^4)$.
\\
\\
      Iterate/cancel/summation method \\

      \begin{align*}
      T(n) &=  \cancel{2T(n/2)} + n^4 \\
      \cancel{2T(n/2)} &=  \cancel{2^2T(n/2^2)} + 2((n/2)^4) \\
      \cancel{2^2T(n/2^2)} &=  \cancel{2^3T(n/2^3)} + 2^2((n/2^2)^4) \\
      \cancel{2^{k-1}T(n/2^{k-1})} &=  \cancel{2^kT(n/2^k)} + 2^{k-1}((n/2^{k-1})^4) \\
      T(n) &= 2^{log_2 n} T(n/2^{log_2 n}) + 2^{(log_2 n) -1}((n/2^{{(log_2 n) -1}})^4) \\
      T(n) &= n T(n/n) + \sum_{i=0}^{(log_2 n)-1}  2^i(n/2^i)^4 \\
      T(n) &= n T(1) + n^4 \sum_{i=0}^{(log_2 n)-1}  2^i/16^i \\
      T(n) &= \Theta( n^4) \\
      \end{align*}
      
    \item Solve this one using the master theorem.
      \[T(n) = 16T(n/4) + n^2\]
      \\
      \begin{eqnarray*}
        f(n) &vs.& n^{log_b a} \\
        n^2  &vs.& n^{log_{4} 16} \\
        n^2  &vs.& n^2 \\
        f(n) &=& \Theta(n^{\log_b a})\\
        T(n) &=& \Theta(n^2\lg n) \\
      \end{eqnarray*}
      
    \item Solve this one using the master theorem.
      \[T(n) = 2T(n/4) + \sqrt{n}\]
      \begin{eqnarray*}
        f(n) &vs.& n^{log_b a} \\
        \sqrt{n}\  &vs.& n^{log_{4} 2} \\
        \sqrt{n}\  &vs.& \sqrt{n}\ \\
        f(n) &=& \Theta(n^{\log_b a})\\
        T(n) &=& \Theta(\sqrt{n}\lg n) \\
      \end{eqnarray*}
    
    \item Solve this one using the master theorem.
      \[T(n) = 4T(n/3) + n\lg n\]
      \begin{eqnarray*}
        f(n) &vs.& n^{log_b a} \\
        n\lg n\  &vs.& n^{log_{3} 4} \\
        f(n) &=& O(n^{\log_b a-\epsilon})\\
        T(n) &=& \Theta(n^{\log_3 4})
      \end{eqnarray*}
    
    \item Solve this one using the master theorem.
      \[T(n) = 4T(n/2) + n^2\sqrt{n}\]
      \begin{eqnarray*}
        f(n) &vs.& n^{log_b a} \\
        n^2\sqrt{n} &vs.& n^{log_2 4} \\
        n^2n^{1/2} &vs.& n^2 \\
        n^{2.5} &vs.& n^2 \\
        f(n) &=& \Omega(n^{log_b a+\epsilon}) \\
        T(n) &=& \Theta(n^{2.5}) \\       
      \end{eqnarray*}
      note:
      $af(n/b) = 4f(n/2) = 4/2^{2.5}f(n) \leq cf(n)$
    
    \end{enumerate}
    
    \end{document}